### 2018-02-11 abstract.tex
did some cleanng heare and there
som edits before submission

% 
It has become common practice to validate ground motion simulation based on a variety of time- and frequency-domain metrics scaled to quantify the level of agreement between synthetics and data, or with respect to other reference solutions. There is, however, no agreement about the importance or weight that it ought to be given to each individual metric. This leads to their selection often being subjective, either based on intended applications or personal preferences. As a consequence, it is difficult for simulators to identify what modeling improvement are needed, which would be potentially easier if they could focused on a reduced number of alternative metrics. We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques used to label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation. In particular, we study the relationships that exist between eleven different metrics. In the process carried out, these metrics are understood as part of a multi-dimensional space. We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects. This allows us to label a given validation dataset into four categories (poor, fair, good, and excellent) following common practice. We then develop two decision trees using a C5.0 algorithm. These decision trees help narrow the number of metrics leading to a validation prediction into the four categories. These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics. Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

% ===========================================================================================
%
% OLD NAEEM VERSION
% 
% The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics-based deterministic and hybrid simulation methods. Unlike for low frequency simulations $(f \le 0.5 Hz)$, for which it has become reasonable to expect a good match between synthetics and data, in the case of high-frequency simulations $(1 Hz \le f)$ it is not possible to match results on a wiggle-by-wiggle basis. This is mostly due to the various complexities and uncertainties involved in earthquake ground motion modeling. Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as means to characterize how the synthetics match the data on qualitative and statistical sense. In general, these metrics provide GOF scores that measure the level of similarity in the time and frequency domains. It is common for these scores to be scaled from 0 to 10, with 10 representing a perfect match. Although using individual metrics for particular applications is considered more adequate, there is no consensus or a unified method to classify the comparison between a set of synthetic and recorded seismograms when the various metrics offer different scores. We study the relationship among these metrics through semi-supervised and supervised learning approaches in order to labeling and classifying data, respectively. In particular, we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics. We put these stations in the category of cannot-link constraints. We generate the dataset through the validation of the results from a deterministic (physics-based) ground motion simulation for a moderate magnitude earthquake in the greater Los Angeles basin using three velocity models. The maximum frequency of the simulation is 4 Hz. The dataset involves over 300 stations and 11 metrics, or features, as they are understood in the machine learning lingo, where the metrics form a multi-dimensional space. We address the high-dimensional feature effects with a subspace-clustering analysis using 2,3, and 4 dimensional spaces. We label the dataset in each subspace as poor, fair, good, and excellent, where, hypothetical stations with score of 3,5,7, and 9 belong them, respectively. We assign the final label based on the mode of subspaces for each station.  Once the dataset has been labeled, we develop two decision trees using C5.0 algorithm as prediction models to classify a simulation in four groups (poor, fair, good, and excellent) based on a reduced number of metrics. We address the unbalanced dataset issue through oversampling method and analyze the precision, recall and F1-score of the classifiers. The prediction models represent a simple, yet effective model to accurately classify the accuracy of the simulation process based on application-independent approach. We analyze the different combinations of simulation accuracy classes for each components and present a guideline for making decision on accuracy level of simulation independent of components. We apply the proposed method on 2008 Chino Hills earthquake data and synthetic and present the results.

>>>>

% 
It has become common practice to validate ground motion simulation based on a variety of time- and frequency-domain metrics scaled to quantify the level of agreement between synthetics and data, or with respect to other reference solutions. There is, however, no agreement about the importance or weight that it ought to be given to each individual metric. This leads to their selection often being subjective, either based on intended applications or personal preferences. As a consequence, it is difficult for simulators to identify what modeling improvement are needed, which would be potentially easier if they could focused on a reduced number of alternative metrics. We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques used to label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation. In particular, we study the relationships that exist between eleven different metrics. In the process carried out, these metrics are understood as part of a multi-dimensional space. We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects. This allows us to label a given validation dataset into four categories (poor, fair, good, and excellent) following common practice. We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories. These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics. Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

### 2018-02-11 conclusions.tex
som edits before submission

We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics. As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset. In particular, the proposed algorithm uses a three disjunctive decisions based on comparisons between data and synthetics in terms of the total energy and acceleration response spectra. We also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies. Similarly, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.

We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulation done for a particular earthquake in a particular region. In a future follow-up study it would be ideal to refine the tree adding additional simulations from different earthquakes, regions, models, etcetera, in order to arrive to a sound and robust decision tree. The procedural steps laid out here, nonetheless, remains valid---and it is likely that the metrics of energy and response spectra will prevail as those among the most decisive ones.

>>>>

We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics, and develop an application independent decision algorithm. As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset. In particular, the proposed algorithm uses three (disjunctive decision) steps based on the values obtained for the metrics of the total energy and acceleration response spectra. we also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies. Similarly, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.

We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake in a particular region. In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, etcetera) in order to arrive to a sufficiently sound and robust decision tree. The procedural steps laid out here, nonetheless, remains valid. We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.

### 2018-02-11 data-resources.tex
som edits before submission

Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018). Additional calculations and figures were done using MATLAB B (\url{http://www.mathworks.com}, last accessed February 2018). Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed February 2018). Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed February 2018). The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.

>>>>

Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018), and the C5.0 package developed by \citet{Kuhn_2017_Manual}. Additional calculations and figures were done using MATLAB (\url{http://www.mathworks.com}, last accessed February 2018). Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed February 2018). Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed February 2018). The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.

### 2018-02-11 dataset.tex
did some cleaning
did some cleanng heare and there
som edits before submission

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations. The color version of this figure is available only in the electronic edition.}
    \label{fig:ref-gof-maps}
\end{figure*}
% 
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD). In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots. The color version of this figure is available only in the electronic edition.}
    \label{fig:data-box-plot}
\end{figure}

Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the component of motions. Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using the whole dataset of scores. The motivation behind this choice was that the dataset, as a sample of GOF values, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}. As such, given the simulations for each velocity model (3), the motion components (3), and the number of stations used in the validation (336) gave us a large enough sample of 3,024 GOF scores. Figure \ref{fig:data-box-plot} illustrates this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components. It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.

% ===========================================================================================
%
% OLD NAEEM VERSION
% 
% State of California is an earthquake prone region with a long history of seismicity. The Los Aneles basin, due to its unique geologic sediments, has been studied from different perspective. Numerous high resolution velocity models are developed mainly to conduct forward ground motion simulation. Occurrence of frequent moderate magnitude earthquakes made the Los Angeles basin as a natural laboratory for seismic  studies specially testing the forward ground motion simulation modeling. In a relatively recent study, \citet{Taborda_2014_BSSA} conducted a comprehensive study in order to assess the functionality of developed velocity models in the Los Angeles basin and surrounding areas. They simulated the 2008 Chino Hills earthquake using 3 velocity models including: CVM-S4, CVM-H, and CVM-H+GTL (for more details about these models see \citet{Taborda_2014_BSSA}.)
% They computed the GOF scores based on \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA} as discussed in the validation section. Fig.~\ref{fig:epicenters} shows the location of the epicenter and seismic stations.

% \begin{figure}
%     \centering
%     \includegraphics
%        % [width=\columnwidth]
%         [width=450px]
%         {figures/pdf/Figure_1}
%     \caption{(a) Region of interest and epicentral location of the 2008 $M_W ~ 5.4$ Chino Hills earthquake (highlighted). Focal mechanism is shown at the margin of the region of interest. Major quaternary faults in the area are shown in the back along with the main roads and county divisions. (b) Simulation area of interest. Dots indicate the location of the 336 stations considered in this study for validation. The main local and interstate roads are also shown here. In both panels, the background shows the hillshade topography of the region. The color version of this figure is available only in the electronic edition.}
%     \label{fig:epicenters}
% \end{figure}

% We use \citet{Taborda_2014_BSSA} database where they computed 11 scores for 3 different velocity models. In this study we only analyze the results for broadband which in this case is \feq{0.1}{4}. Fig.~\ref{fig:data_box_plot} shows the box plot of all data that we used in this study which we separate them in velocity models and components. In total we use $336 \times 3 \times 3 = 3024$ individual pair of data and synthetic.

% \begin{figure}
%     \centering
%     \includegraphics
%        % [width=\columnwidth]
%         [width=\textwidth]
%         {figures/pdf/Figure_2.pdf}
%     \caption{Box plot of data used in the study. Metrics are shown for C1 to C11 for 3 components and in separate plots for velocity models. Median of data are shown as a notch. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
%     \label{fig:data_box_plot}
% \end{figure}

% As we explained before, we use all data set regardless of velocity models and components. However, distinguishing data for velocity model and components can answer the question that if the goodness of fit scores dependent on them. Addressing this research question is beyond the scope of this paper. We distinguish data based on velocity models and components only in presenting data or results.

% Fig.~\ref{fig:data_density}  shows the density distribution of each score for CVM-S. The figure shows the variation of different metrics with components. Except in cross correlation (C10) in all metrics the up-down (UD) component scores are about the same or higher than horizontal components.

% \begin{figure}
%     \centering
%     \includegraphics
%        % [width=\columnwidth]
%         [width=\textwidth]
%         {figures/pdf/Figure_3.pdf}
%     \caption{Distribution of data (only CVM-S) for different components in different scores based on different components. Each distribution has a unit area.}
%     \label{fig:data_density}
% \end{figure}

>>>>

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations. The color version of this figure is available only in the electronic edition.}
    \label{fig:ref-gof-maps}
\end{figure*}
% 
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD). In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots. The color version of this figure is available only in the electronic edition.}
    \label{fig:data-box-plot}
\end{figure}

Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the component of motions. Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using the whole dataset of scores. The motivation behind this choice was that the dataset, as a sample of GOF values, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}. As such, given the simulations for each velocity model (3), the motion components (3), and the number of stations used in the validation (336) gave us a large enough sample of 3,024 GOF scores. Figure \ref{fig:data-box-plot} illustrates this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components. It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component. Consequently, we combine all the data into a single dataset.

### 2018-02-11 introduction.tex
did some cleanng heare and there
som edits before submission

To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus the strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results. We treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques. We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent. These categories adhere to the common practice in validation. The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}. Upon labeling the dataset, we develop two decision trees using a C4.5 algorithm \citep[][]{Quinlan_1993_Book} to prioritize the metrics involved in the validation process. The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.

@@

% ===========================================================================================
%
% OLD NAEEM VERSION
% 
% The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics based deterministic and hybrid simulation methods. Unless in very low frequencies (\fleq{0.5}), due to limitations and uncertainties in the ground motion simulation and velocity models, it is not possible to match the simulation and synthetics wiggle by wiggle. Therefor, in order to compare synthetic seismogram with recorded data as a complex time series, different metrics are defined. These metrics are means to characterize the how well the synthetic matches the statistical characteristic of observed data. In general, these metrics compute the goodness-of-fit (GOF) between data and synthetic in time and frequency domains. \citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterize the GOF in different frequency bands. \citet{Kristekova_2006_BSSA,Kristekova_2009_GJI} developed envelope and phase misfit criteria as well as GOF metrics for realistic scenarios. \citet{Olsen_2010_SRL} presented an alternative GOF metrics for broadband synthetics with relatively high resolution of the small misfits, and \citet{Taborda_2013_BSSA} added modification to \citet{Anderson_2004_Proc} metrics.

% These metrics have been widely used in different studies. Recently, due to advances in computational resources, 3D ground motion simulation on a regional scale were conducted in different regions using different methods \citep[e.g.,][]{Taborda_2016_GJI,Komatitsch_2004_BSSA,Bielak_2010_GJI,Roten_2016_GJI}. The accuracy of simulation with observed data are measured using these GOF scores. However, researchers use these metrics in different perspectives which is highly dependent on their application. For example, for structural engineers who are interested in designing tall buildings, among other metrics, displacement could be an important factor. However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important criteria. In general, for all purposes, the duration of strong ground morion could be an important factor. Analysis of the strong ground motion simulation for a specific application based on individual metrics is an acceptable practice, however, validating the simulation process only with individual score is not a valid approach. Previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}. \citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores presenting a general overview of GOF. However, the metric is a linear combination and somehow is biased by the authors' preferences about the metrics.

% Physics-based ground motion simulation communities push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations. Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, there is no consensus on the most important metrics, specially when they have significantly different GOF score. Therefore, the question "How one can estimate the accuracy of simulation based on all scores or which metrics can be good representatives of simulation accuracy on behalf of other metrics?" has not addressed. This research question motivated us to approach the problem from different perspective. In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations or which metrics are the best representer in classifying the stations. However, as far as the authors knowledge, there is no metric to be able to classify a pair of stations based on GOF scores as an estimation of the overall accuracy of the simulation. Therefore, as a first step, if we label the pair of station for simulation accuracy we can discuss the relationships between GOF scores. To do this we need a dataset where involves GOF scores for a good amount of stations.

% \citet{Taborda_2014_BSSA} conducted a comprehensive study to analyze the functionality of different velocity models in the Los Angeles basin. They used 2008 Chino Hills earthquakes based on 3 velocity models and conducted simulations with maximum frequency of \fmaxeq{4}. They compared the accuracy of the velocity models through using the GOF scores. The study generated an invaluable synthetic and data  GOF scores dataset.\\
% The first and the most common method in categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}. However, there is a drawback in this method. If the data is not explicitly distinguishable the final clusters are highly dependent on initial starting points of the algorithm. Therefor, for stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis. The main reason for this issue is the algorithm solely looks for data which are close together in terms of Euclidean distance. However, in our case we have one more extra information. Hypothetically, if we have a pair of synthetic and data with GOF score equal 3 for all metrics, according to \citet{Anderson_2004_Proc} we can classify the station as poor with a strong confidence. This is also true for GOF score of 5, 7, and 9 for fair, good, and excellent classes, respectively. This background knowledge about the database could give a strong formation to the clustering process. \citet{Wagstaff_2001_Proc} defined a \kmeans{} clustering using background knowledge. Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial clusters' centers. Using this approach alongside the subspace analysis (more details are presented in the section.~\ref{clustering}) we cluster the stations, then in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores. We tune the learning algorithm and present the functionality of each prediction model based on different metrics. It is possible to have different classes for different components of a pair if synthetic and data. We study the different combination of components' classes in different distances and propose a rough estimate on deciding about the final GOF class of a pair of synthetic and data which is application and component independent.\\
% In the following section,we present a summary of the validation metrics then we discuss the dataset. Later on we explain the methodology. We provide a brief summary of unsupervised  and semi-supervised learning which are ordinary and constrained clustering approaches, respectively. We also discuss the basics of decision tree algorithm as a supervised learning method. We discuss the clustering challenges and use subspace analysis in the clustering analysis section. We also discuss the imbalanced data effects and solutions in the classification model section. We develop two decision tree classifiers (prediction model) based on labeled data for future use in comparing data and synthetics. We propose an algorithm to give a rough estimate to have component independent prediction. We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset and present the spatial variation of the simulation goodness of fit classes on the Los Angles basin for three velocity models and three components in component dependent and for three velocity models in component independent approaches.

>>>>

To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus the strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results. We treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques. We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent. These categories adhere to the common practice in validation. The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}. Upon labeling the dataset, we develop a family of decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}, and select a few to prioritize the metrics involved in the validation process. The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.

@@
